analysis

### 2020-05-29 analysis.Rmd
refined GAM analysis

### Prepare environment
```{r setup, include=FALSE}
library(bipartite)
library(lme4)
library(lmerTest)
library(tidyverse)
library(lubridate)
library(GGally)
```

@@

@@

>>>>

### Prepare environment
```{r setup, include=FALSE}
library(bipartite)
library(mgcv)
library(mgcViz)
library(lme4)
library(lmerTest)
library(tidyverse)
library(lubridate)
library(GGally)
```

@@

### Web asymmetry ~ gdd
ggplot(abund_div, aes(gdd.cum, web.asymmetry)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  facet_grid(year ~ phase, scales = "free")

### Web asymmetry ~ yday
ggplot(abund_div, aes(yday, web.asymmetry)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  facet_grid(year ~ phase, scales = "free")

@@

### GAM surface plots of abundance and diversity metrics (just for fun)
#### 2012
```{r}
### Per-capita resource availability
percap_yday <- gam(log(fl.per.bb) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
                   data = filter(abund_div, year == 2012),
                   method = "REML")

percap_yday.vis <- getViz(percap_yday)
check.gamViz(percap_yday.vis)
plot(sm(percap_yday.vis, 1)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("Per-capita FL (yday)") +
  xlab("Day of year") +
  ylab("Elevation")

percap_gdd <- gam(log(fl.per.bb) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
                  data = filter(abund_div, year == 2012),
                  method = "REML")

percap_gdd.vis <- getViz(percap_gdd)
check.gamViz(percap_gdd.vis)
plot(sm(percap_gdd.vis, 1)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("Per-capita FL (GDD)") +
  xlab("Day of year") +
  ylab("Elevation")

### Web asymmetry
#Create a version of abund_div in which web.asymmetry has been forced into postive values (to enable transformation)
abund_div_trans <- abund_div %>%
  mutate(web.asymmetry = web.asymmetry - min(web.asymmetry - 0.1))

ggplot(abund_div_trans, aes(web.asymmetry)) +
  geom_histogram()

webass_yday <- gam(log(web.asymmetry) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
                   data = filter(abund_div_trans, year == 2012),
                   method = "REML")

webass_yday.vis <- getViz(webass_yday)
check.gamViz(webass_yday.vis)
plot(sm(webass_yday.vis, 1)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("Web asymmetry (yday)") +
  xlab("Day of year") +
  ylab("Elevation")

webass_gdd <- gam(log(web.asymmetry) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
                  data = filter(abund_div_trans, year == 2012),
                  method = "REML")

webass_gdd.vis <- getViz(webass_gdd)
check.gamViz(webass_gdd.vis)
plot(sm(webass_gdd.vis, 1)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("Web asymmetry (GDD)") +
  xlab("Day of year") +
  ylab("Elevation")

### Floral abundance
fl_abund_yday <- gam(log(fl.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
                    data = filter(abund_div, year == 2012),
                    method = "REML")

fl_abund_yday.vis <- getViz(fl_abund_yday)
check.gamViz(fl_abund_yday.vis)
plot(sm(fl_abund_yday.vis, 1)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("FL abundance (yday)") +
  xlab("Day of year") +
  ylab("Elevation")

fl_abund_gdd <- gam(log(fl.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
                    data = filter(abund_div, year == 2012),
                    method = "REML")

fl_abund_gdd.vis <- getViz(fl_abund_gdd)
check.gamViz(fl_abund_gdd.vis)
plot(sm(fl_abund_gdd.vis, 1)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("FL abundance (GDD)") +
  xlab("Day of year") +
  ylab("Elevation")

### BB abundance
bb_abund_yday <- gam(log(bb.abund) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
                    data = filter(abund_div, year == 2012),
                    method = "REML")

bb_abund_yday.vis <- getViz(bb_abund_yday)
check.gamViz(bb_abund_yday.vis)
plot(sm(bb_abund_yday.vis, 1)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("BB abundance (yday)") +
  xlab("Day of year") +
  ylab("Elevation")

bb_abund_gdd <- gam(log(bb.abund) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
                    data = filter(abund_div, year == 2012),
                    method = "REML")

bb_abund_gdd.vis <- getViz(bb_abund_gdd)
check.gamViz(bb_abund_gdd.vis)
plot(sm(bb_abund_gdd.vis, 1)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("BB abundance (GDD)") +
  xlab("Day of year") +
  ylab("Elevation")

### BB richness
bb_rich_yday <- gam(log(bb.rich) ~ te(yday, elev.mean, bs = "tp", m = 2, k = 10),
                    data = filter(abund_div, year == 2012),
                    method = "REML")

bb_rich_yday.vis <- getViz(bb_rich_yday)
check.gamViz(bb_rich_yday.vis)
plot(sm(bb_rich_yday.vis, 1)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("BB richness (yday)") +
  xlab("Day of year") +
  ylab("Elevation")

bb_rich_gdd <- gam(log(bb.rich) ~ te(gdd.cum, elev.mean, bs = "tp", m = 2, k = 10),
                    data = filter(abund_div, year == 2012),
                    method = "REML")

bb_rich_gdd.vis <- getViz(bb_rich_gdd)
check.gamViz(bb_rich_gdd.vis)
plot(sm(bb_rich_gdd.vis, 1)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("BB richness (GDD)") +
  xlab("Day of year") +
  ylab("Elevation")
```
